### 2012-01-04 GUIErrorDetector/paper/technique.tex
address Mike's comments on the technique section

We first gives our high-level approach formulation, then present
an error detection algorithm. Finally, we show how to instantiate
the error algorithm to different GUI frameworks.

@@

This section formulates our high-level approach. We define the
notations of \textit{UI thread}, \textit{non-UI threads},
\textit{UI accessing methods},
 \textit{Safe methods}, and \textit{invalid thread access error}, and
two assumptions we make with regard to error detection.

\noindent {\textsc{\textbf{Definition 1 (UI Thread).}}} {UI thread
is a special thread created by the GUI framework during
GUI initialization. After the GUI becomes visible, the UI thread
takes charge of the application to handle events from the GUIs,
or spawn new threads to process lengthy operations in the background. }\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 1.}}} {We assume that each multithreaded
GUI application has a single global UI thread. This is true for applications
built on top of GUI frameworks adapting the single-thread rule. The
only exception is that 
 an application may folk a new process to launch another
application with its own UI thread. In that case, we require the new
launched application to be analyzed separately.}\vspace{1mm}

@@

\noindent {\textsc{\textbf{Assumption 2.}}} { We assume that each non-UI
thread can only be spawned by the UI thread. Under this assumption,
 all post-initialization
GUI work naturally occurs in the UI-thread. Once the GUI is visible, the
application is driven by events, which are always handled in the UI thread.
We can safely ignore non-UI threads that are spawned during pre-initilaizaiton
GUI work, since if such threads access a GUI object, an exception becomes
immediately apparent and the whole application may abort even before the
GUI is visible. This is highly unlikely for most of the realistic GUI
applications.
}\vspace{1mm}

\noindent {\textsc{\textbf{Definition 3 (UI Accessing Methods).}}} { A method
whose execution may read or write a UI object is called a UI accessing method.}\vspace{1mm}

\noindent  {\textsc{\textbf{Definition 4 (Safe Methods).}}} {GUI frameworks that
adapt the single-thread rule must provide methods to permit non-UI threads
to run code in the UI thread. We can such methods \textit{safe methods}, since
they can be invoked safely by any threads.}\vspace{1mm}

\noindent  {\textsc{\textbf{Definition 5 (Invalid Thread Access).}}} {An invalid
thread access error occurs when the UI thread spawns a non-UI thread, and there
exists a path from the non-UI thread's \CodeIn{start} method to any UI accessing methods
without going through any safe methods. }\vspace{2mm}

@@

The algorithm for detecting potential invalid thread access errors
is shown in Figure~\ref{fig:detectalgorithm}. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specially, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretical ideal call graph can be defined as the union of the
dynamic call graphs over all possible executions of the program. 
A conservative static call graph used in practice is a superset of
the ideal call graph; it over-approximates the
dynamic call graph of every possible execution. 
In an object-oriented programming language like
Java, a method is the basic logic unit in a program. Reporting
potential errors at the method level can be sufficient in most cases.
Thus, we believe using call graph is well suited for this problem while
compared with other
program representations like control flow graph or program dependence
graph.

Our algorithm in Figure~\ref{fig:detectalgorithm} works as follows. It
first constructs a static call graph for the tested program (line 2),
then specifies entry nodes for it (line 3). The algorithm starts to traverse
the call graph from every entry node (line 6). For every entry node,
it first uses Breadth First Search (BFS) to find all reachable
\CodeIn{Thread.start()} method (line 7). Each \CodeIn{Thread.start()} method
indicates that a new, non-UI thread is spawned.  After that, 
the algorithm uses a slightly modified 
BFS algorithm to find reachable UI accessing methods (lines 9 -- 23)
from each \CodeIn{Thread.start()} node.
Specifially, it uses a queue to keep track of all visited nodes, and
then checks every successive node to be visited (line 12);
it reports an error if that node
is a UI accessing method (line 15), or skips it if it is
a safe method (line 17), or simply keeps traversing otherwise (line 19).

@@

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph for partial programs such as an eclipse plugin,
or in the presence of reflection, is non-trivial. 
In particular, GUI applications built on some popular frameworks
often use configuration files and reflection to specify GUI layout
or extension points.
Existing call graph construction algorithms either suffer from incompleteness
or imprecision.  To alleviate this problem, 
We proposed a combined RTA
and k-CFA call graph construction algorithm in Section~\ref{sec:cg}.

We also note that a static analysis may report false positives due
to its conservative nature, or multiple warnings that acutally corrspond to
the same error. In addition, some GUI applications may use
Java native methods to interact with the underlying operating system.
Native methods are often beyond the ability of a static analysis, but
should be considered to construct a good call graph. To address the
above two limitations, we present a set of heuristics to filter
highly likely false positives and remove redundant reports, and provide
annotation support for programmers to specify call relationship for
 native methods.

@@

1. public class MyActivity extends Activity \{
2.    final List<View> cachedViews = new LinkedList<View>();
3.    @Override
4.    public void onCreate(Bundle savedInstanceState) \{
5.        super.onCreate(savedInstanceState);
6.        setContentView(R.layout.main);
7.        Button button = (Button) findViewById(R.id.\textbf{button\_id});
8.        button.setOnClickListener(new Button.OnClickListener() \{
9.            @Override
10.           public void onClick(View v) \{
11.               button.setText("Button Clicked.");
12.               cachedViews.add(v);
13.           \}
14.       \});
15.   \}
16. \}
\end{alltt}
\end{CodeOut}
\label{fig:sampleandroid}
\caption{Sample GUI application code on the Android platform. The layout
XML file (the top) specifies a Button object declaratively, 
and the Java code (the bottom)
first loads the XML file (line 6) and then uses reflection to create
a Button object by its ID (line 7, the \CodeIn{findViewByd} method).}
\end{figure}

@@

Unfortunately, GUI applications often use reflection, which makes either
RTA or k-CFA insufficient to build a satisfactory call graph. 
The example code in Figure~\ref{fig:sampleandroid}  illustrates their limitations.
In Figure~\ref{fig:sampleandroid}, line 7 uses reflection to create a \CodeIn{Button}
object by looking up its id declared in the XML file. When this button
is clicked, its event handling code (line 10 -- 13) updates the text and cached
the current view to a list declared at line 2.

>>>>

We first give our high-level approach formulation, then present
an error detection algorithm. Finally, we show how to instantiate
the error algorithm to different GUI frameworks.

@@

This section formulates our high-level approach. We define the
notations of \textit{UI thread}, \textit{non-UI threads},
\textit{UI accessing methods},
 \textit{safe methods}, and \textit{invalid thread access error}, and
two assumptions we make with regard to error detection.

\noindent {\textsc{\textbf{Definition 1 (UI Thread).}}} {The UI thread
is a special thread created by the GUI framework during
GUI initialization. After the GUI becomes visible, the UI thread
takes charge of the application to handle events from the GUI,
and spawns new threads to process lengthy operations in the background. }\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 1.}}} {We assume that each multithreaded
GUI application has a single global UI thread. This is true for applications
built on top of GUI frameworks adapting the single-thread rule. The
only exception is that 
 an application may fork a new process to launch another
application with its own UI thread. In that case, we require the new
launched application to be analyzed separately.}\vspace{1mm}

@@

\noindent {\textsc{\textbf{Assumption 2.}}} { We assume that each non-UI
thread is spawned by the UI thread. Under this assumption,
we ignore all non-UI threads created by the GUI framework
before the UI thread has been initialized. That is, all post-initialization
GUI work naturally occurs in the UI thread. Once the GUI is visible, the
application is driven by events, which are always handled in the UI thread.
We believe this assumption is reasonable, since if a non-UI thread 
spawned during pre-initialization GUI work accesses a GUI object, an exception becomes
immediately apparent and the whole application may abort even before the
GUI is visible. This is highly unlikely for realistic GUI
applications.
}\vspace{1mm}

\noindent {\textsc{\textbf{Definition 3 (UI-Accessing Methods).}}} { A method
whose execution may read or write a UI object is called a UI accessing method.}\vspace{1mm}

\noindent  {\textsc{\textbf{Definition 4 (Safe Methods).}}} {GUI frameworks that
adopt the single-thread rule must provide methods to permit non-UI threads
to run code in the UI thread. We can such methods \textit{safe methods}, since
they can be invoked safely by any thread.}\vspace{1mm}

\noindent  {\textsc{\textbf{Definition 5 (Invalid Thread Access).}}} {An invalid
thread access error occurs when the UI thread spawns a non-UI thread, and there
exists a path from the non-UI thread's \CodeIn{start} method to any UI-accessing method
without going through any safe method. }\vspace{2mm}

@@

The algorithm for detecting potential invalid thread access errors
is shown in Figure~\ref{fig:detectalgorithm}. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specially, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretically ideal call graph can be defined as the union of the
dynamic call graphs over all possible executions of the program. 
A conservative static call graph used in practice is a superset of
the ideal call graph; it over-approximates the
dynamic call graph of every possible execution. 
In an object-oriented programming language like
Java, a method is the basic logic unit in a program. Reporting
potential errors at the method level can be sufficient in most cases.
Thus, we believe using the call graph is well suited for this problem while
compared with other program representations like control flow graph or
program dependence graph.

Our algorithm in Figure~\ref{fig:detectalgorithm} works as follows. It
first constructs a static call graph for the tested program (line 2),
then specifies entry nodes for it (line 3). The algorithm starts to traverse
the call graph from every entry node (line 6). For every entry node,
it first uses Breadth First Search (BFS) to find all reachable
\CodeIn{Thread.start()} calls (line 7). Each \CodeIn{Thread.start()} call
indicates that a new, non-UI thread is spawned.  After that, 
the algorithm uses a BFS to find reachable UI accessing methods (lines 9 -- 23)
from each \CodeIn{Thread.start()} node.
Specifially, it uses a queue to keep track of all visited nodes, and
then checks every successive node to be visited (line 12);
it reports an error if that node
is a UI accessing method (line 15), or skips it if it is
a safe method (line 17), or simply keeps traversing otherwise (line 19).

@@

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph for partial programs such as an Eclipse plugin,
or in the presence of reflection, is non-trivial. 
In particular, GUI applications built on some popular frameworks
often use configuration files and reflection to specify GUI layout
or extension points.
Existing call graph construction algorithms suffer from incompleteness
, imprecision, or both.  To alleviate this problem, 
we proposed a combined RTA
and k-CFA call graph construction algorithm in Section~\ref{sec:cg}.

We also note that a static analysis may report false positives due
to its conservative nature, or multiple warnings that actually correspond to
the same error. In addition, some GUI applications may use
Java native methods to interact with the underlying operating system.
Native methods are often beyond the ability of a static analysis, but
should be considered to construct a good call graph. To address the
above two limitations, we present a set of heuristics to filter
highly likely false positives and remove redundant reports, and provide
annotation support for programmers to specify call relationships for
 native methods.

@@

1. public class MyActivity extends Activity \{
2.    final List<View> cachedViews = new LinkedList<View>();
3.    @Override
4.    public void onCreate(Bundle savedInstanceState) \{
5.        super.onCreate(savedInstanceState);
6.        setContentView(R.layout.main);
7.        Button button = (Button) findViewById(R.id.\textbf{button\_id});
8.        button.setOnClickListener(new Button.OnClickListener() \{
9.            @Override
10.           public void onClick(View v) \{
11.               button.setText("Button Clicked.");
12.               cachedViews.add(v);
13.           \}
14.       \});
15.   \}
16. \}
\end{alltt}
\end{CodeOut}
\label{fig:sampleandroid}
\caption{Sample GUI application code on the Android platform. The layout
XML file (the top) specifies a Button object declaratively, 
and the Java code (the bottom)
first loads the XML file (line 6) and then uses reflection to create
a Button object by its ID (line 7, the \CodeIn{findViewById} method).}
\end{figure}

@@

Unfortunately, GUI applications often use reflection, which makes either
RTA or k-CFA insufficient to build a satisfactory call graph. 
The example code in Figure~\ref{fig:sampleandroid}  illustrates their limitations.
In Figure~\ref{fig:sampleandroid}, line 7 uses reflection to create a \CodeIn{Button}
object by looking up its id declared in the XML file. When this button
is clicked, its event handling code (line 10 -- 13) updates the text and caches
the current view to a list declared at line 2.
